---
title: "An Analysis of Political Affiliations and Respective Environmental Beliefs"
author: "Eva Wilson"
date: "11/22/2021"
output: 
  html_document: 
    theme: readable
    code_folding: hide
---

- Consider creating a final table 
- Add one more sentence to analysis

```{r setup, include = FALSE, message = FALSE, warning = FALSE }
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages:
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)
library(equatiomatic)
```

### 1. Overview 

In the following document, we describe the relationship between individuals' political affiliations and whether or not they consider themselves an environmentalist. 

The data are from a survey conducted by Chelsea Batavia and Andrews Forest LTER to explore "Influences on charitable giving for conservation." The survey was conducted in August 2017 and yielded 1,331 individual respondents.

```{r include = FALSE} 
# read in the data:
survey <- read.csv(here("data","conservation_survey.csv")) %>% 
  clean_names()
```

```{r include = FALSE}
# wrangle the data:

# select variables env_id and pol_or
# add a new column that bins the political party into "conservative" and "liberal"
# exclude political affiliations (4,8,9)

env_pol <- survey %>% 
  select(env_id, pol_or) %>% 
  mutate(pol_affiliation = case_when(
    pol_or %in% c(1,2,3) ~ "conservative",
    pol_or %in% c(5,6,7) ~ "liberal")) %>% 
  mutate(environmentalist = case_when(
    env_id %in% c(1) ~ "Yes",
    env_id %in% c(2) ~ "No",
    env_id %in% c(3) ~ "Unsure")) %>% 
  drop_na(pol_affiliation)
```

#### Create a table that shows the relationship between political affiliation (liberal/conservative) and identification as an environmentalist

```{r}
env_pol %>% 
  count(pol_affiliation, environmentalist)

env_pol_counts <- env_pol %>% 
  janitor::tabyl(pol_affiliation, environmentalist)

env_pol_proportions <- env_pol_counts %>%
  janitor::adorn_percentages() %>%
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
```

```{r include = FALSE}
# create a contingency table
env_pol_ct <- env_pol_counts %>% 
  column_to_rownames(var = "pol_affiliation")

env_pol_ct
```

```{r include = FALSE}
# conduct a chi-square test
env_pol_x2 <- chisq.test(env_pol_ct)

env_pol_x2

env_pol_tidy <- broom::tidy(env_pol_x2)

env_pol_tidy
```

There is a significant association (i.e. non-independence) between an individual's political affiliation and their identification as an environmentalist ($\chi$^2^(`r env_pol_tidy$parameter`) = `r env_pol_tidy$statistic`, *p* = `r env_pol_tidy$p.value`). 

### References

**Data citation**: Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c

Metadata, including survey questions and possible responses, are here: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-and.5444.3
